---
title: "P8105 HW1"
author: 'Yutong Mao (UNI: ym3139)'
date: "2025-09-20"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Problem 1
```{r p1-setup, message=FALSE, warning=FALSE}
library(moderndive)
library(ggplot2)

data("early_january_weather")  
wx <- early_january_weather

```
The dataset `early_january_weather` has `r nrow(wx)` rows and `r ncol(wx)` columns.  
Important variables include: `r paste(names(wx), collapse = ", ")`.  
The average temperature is `r round(mean(wx$temp, na.rm = TRUE), 2)` °F.

```{r}
p_temp_time <- ggplot(wx, aes(x = time_hour, y = temp, color = humid)) +
  geom_point(alpha = 0.55, size = 1.1) +
  labs(
    x = "Time (hourly)",
    y = "Temperature (°F)",
    color = "Humidity",
    title = "Hourly temperature in early January (colored by humidity)"
  ) +
  theme_minimal()

p_temp_time

```

The figure shows: Temperature fluctuates significantly throughout the day (higher during the day, lower at night); and the color indicates a trend of inverse correlation between humidity and temperature (periods of lower temperature are often accompanied by higher humidity).

```{r}
ggplot2::ggsave(
  filename = "fig_temp_vs_time.png",
  plot = p_temp_time,
  width = 7, height = 4, dpi = 300
)
```

